---
title: 'Regression Models: Assignment 1'
author: "Daniel Alonso"
date: "November 24th, 2020"
output: 'pdf_document'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing libraries

```{r, echo=TRUE, warning=FALSE, message=FALSE, results="hide"}
library(dplyr)
library(MuMIn)
```

# Exercise 1

## Simulation

```{r, echo=TRUE, warning=FALSE, results='hide'}
sim = list()
for (j in 1:1000) {
    vals = c()
    for (i in 1:100) {
        run = 3 + 3*cos(i/10 + 50) + rnorm(1, mean=0, sd=1)
        vals = c(vals, run)
    }
    sim[[j]] = vals
}
sim
```

```{r, echo=TRUE, warning=FALSE}
```

# Exercise 2

## Importing the data

```{r, echo=TRUE, warning=FALSE}
d <- data.frame(read.table('../data/index.txt', header=TRUE))
```

```{r, echo=TRUE, warning=FALSE}
X = d$PovPct
Y = d$Brth15to17
beta1 = cov(X, Y)/var(X)
beta0 = mean(Y) - beta1*mean(X)
beta1
beta0
```

\newpage

# Exercise 3

First we have the log-likelihood function for $\beta$ and $\sigma^{2}$

$l(\sigma^{2} | X) = \sum_{i=1}^n log(\frac{1}{\sqrt{2 \pi \sigma^{2}}} - \frac{(Y_{i} - (\beta_{0} + \beta_{1} x_{ik} + \dots + \beta_{k} x_{ik}))^{2}}{2 \sigma^{2}})$

$\propto - \frac{n}{2} log(\sigma^{2}) - \frac{(Y - X \beta) \prime (Y - X \beta)}{2 \sigma^{2}}$

Differentiating the second expression:

$\frac{\partial l}{\partial \sigma} ( - \frac{n}{2} log(\sigma^{2}) - \frac{(Y - X \beta) \prime (Y - X \beta)}{2 \sigma^{2}} = 0$

We get:

$- \frac{n}{2} (\frac{1}{ \sigma^{2}} ) (2 \sigma) - (Y - X \beta) \prime (Y - X \beta) * (-2)(2 \sigma^{-3}) = 0$

We reduce the expression further:

$- \frac{n}{\sigma} +  \frac{(Y - X \beta) \prime (Y - X \beta)}{\sigma^{3}} = 0$

We multiply both sides by $\sigma^{3}$ and we get:

$- n \sigma^{2} +  (Y - X \beta) \prime (Y - X \beta) = 0$

And solving for $\sigma^{2}$ we get:

$\hat{\sigma^{2}} = \frac{(Y - X \beta) \prime (Y - X \beta)}{n}$

Which is our maximum likelihood estimator for $\sigma^2$

# Exercise 4

# Exercise 5

```{r, echo=TRUE, warning=FALSE}
bodyfat <- data.frame(read.table('../data/bodyfat.txt', header=TRUE))
modall <- lm(hwfat ~., data = bodyfat)
summary(modall)
```

The sum of residuals is zero:

```{r, echo=TRUE, warning=FALSE}
residuals <- sum(resid(modall))
```

The sum of the observed data is equal to the sum of the fitted values

```{r, echo=TRUE, warning=FALSE}
Y_hat <- predict(modall, bodyfat[1:length(names(bodyfat))-1])
sum(bodyfat$hwfat) - sum(Y_hat)
```

The residuals are orthogonal to the predictors

```{r, echo=TRUE, warning=FALSE}
sum(residuals*bodyfat[1:length(names(bodyfat))-1])
```

The residuals are orthogonal to the fitted values

```{r, echo=TRUE, warning=FALSE}
sum(residuals*Y_hat)
```

# Exercise 6

```{r, echo=TRUE, warning=FALSE}
# rsq <- function(x,y) cor(x,y)^2
# cols <- names(bodyfat)[1:length(names(bodyfat))-1]
# r_2 <- c()
# names(r_2) <- cols
# for (i in 1:length(cols)) {
#     modall <- lm(hwfat ~ cols[i], bodyfat)
#     r_2 <- rsq(predict(hwfat))
# }
# r2
```

```{r, echo=TRUE, warning=FALSE}
options(na.action = "na.fail")
modall <- lm(hwfat ~., data = bodyfat)
combs <- dredge(modall, extra = "R^2")
print("best model")
combs[combs$"R^2" == max(combs$"R^2")]
```

